---
title: "2021_abundance_demo"
author: "Omar Al-Saleem"
date: "10/20/2021"
output: github_document
---
# Goal 

This document shows how **individual bottle** bacterial abundance data from 2021 remineralization experiments were processed, QC'd and analyzed. It also provides an intro to data processing and analysis with Rstudio and R markdown.

#Intro to R markdown
Load packages that we'll need to analyze our data.
```{r Load Packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(lubridate)
```

#Import data

```{r}
excel_sheets("~/Desktop/github/144l_students_2021/Input_Data/week4/144L_2021_BactAbund.xlsx")

metadata <- read_excel("~/Desktop/github/144l_students_2021/Input_Data/week4/144L_2021_BactAbund.xlsx", sheet="Metadata")

glimpse(metadata)
#unique(metadata$Bottle)
#unique(metadata$Treatment)

data <- read_excel("~/Desktop/github/144l_students_2021/Input_Data/week4/144L_2021_BactAbund.xlsx", sheet="Data")

joined<- left_join(metadata, data)#attach data to metadata which will join the right dataset to the left one by using variables that are the same across the two dataframes
glimpse(joined)
```

#Prepare The Data

we will convert the Data and Time values from characters to dates by adding columns with the time elapsed for each treatment, and convert to cells/L because it will match up with the TOC data later. Also, we will sbset the data for variables of interest and drop the NA values.

to do this, we will use the pipe operator "%>%", which basically says like "first do one thing to the data. THEN, do this other thing." (with the %>% operator taking the place of the word THEN in this scenario). Every call that we invoke with an additional pipe is going to take place on the variable (dataframe) that we specify at the beginning.

```{r message=FALSE, warning=FALSE}
cells <- joined %>%
  mutate(Datetime= ymd_hm(Datetime), #splits apart Datetime as specified
  all_cells_L = as.numeric(all_cells_uL) *1000000) %>%
  group_by(Treatment, Bottle) %>%
  mutate(interv=interval(first(Datetime), Datetime),
         s = as.numeric(interv),
         hours = s/3600,
         days = hours/24) %>%
  ungroup() %>%
  select(Experiment:DNA_Sample, all_cells_L, hours, days) %>%
  drop_na(all_cells_L)
glimpse(cells)
 
```
# Plot Growth Curves

we will plot growth curve for each bottle. Need cell abundance and days data.

```{r}
#assign hex colors for different treatments
custom.colors <- c("Control" = "#377EB8", "Kelp Exudate" = "#4DAF4A", "Kelp Exudate_Nitrate_Phosphate"= "#E41A1C", "Glucose_Nitrate_Phosphate" = "#FF7F00")

#assign levels to control what order things will appear in the legend
levels <-c("Control", "Kelp Exudate", "Kelp Exudate_Nitrate_Phosphate", "Glucose_Nitrate_Phosphate")

#now will use a handy package, ggplot to visualize the data

cells %>%
  mutate(dna = ifelse(DNA_Sample == T, "*", NA)) %>%
  ggplot(aes(x=days, y=all_cells_L, group = interaction(Treatment, Bottle))) +
  geom_line(aes(color = factor(Treatment, levels = levels)), size=1) +
  geom_point(aes(fill = factor(Treatment, levels = levels)), size = 3, color= "black", shape = 21) +
  geom_text(aes(label = dna), size = 12, color = "#E41A1C") +
  labs(x = "Days", y = expression(paste("Cells, L"^-1)), fill = "")+
  guides(color = "none") +
  scale_color_manual(values = custom.colors) +
  scale_fill_manual(values = custom.colors)
  #facet_grid(rows = "Treatment")
  theme_bw()
glimpse(cells)
```


# Next Steps

We can calculate:

-total change in cells from the initial condition to the end of the experiment
-specific growth rate as the slope of ln(abundance) v time during exponential growth phase
-doubling time as ln(2) divided by the specific growth rate 
-mean of each of these parameters across each treatment

first, we need to determine **where** exponential growth is occurring in each of the bottles if it happens. To do this, we 'll plot ln(abundance) vs time.

# Identify exponential phase of growth in our remin experiments

```{r}
ln_cells <- cells %>%
  group_by(Treatment, Bottle) %>%
  mutate(ln_cells = log(all_cells_L), 
         diff_ln_cells = ln_cells - lag(ln_cells, default = first(ln_cells)))
```

Now, lets plot our newly calcuated data!!

```{r fig.height=6, fig.width=8}
ln_cells %>%
  mutate(dna = ifelse(DNA_Sample == T, "*", NA)) %>%
  ggplot(aes(x=days, y=diff_ln_cells, group = interaction(Treatment, Bottle))) +
  geom_line(aes(color = factor(Treatment, levels = levels)), size=1) +
  geom_point(aes(fill = factor(Treatment, levels = levels)), size = 3, color= "black", shape = 21) +
  geom_text(aes(label = dna), size = 12, color = "#E41A1C") +
  labs(x = "Days", y = expression(paste("∆ln cells, L"^-1)), fill = "")+
  guides(color = "none") +
  scale_color_manual(values = custom.colors) +
  scale_fill_manual(values = custom.colors) +
  facet_wrap("Bottle" , ncol =2) +
  theme_bw()
```

Exponential growth seems to be occuring at the begging of the expierment between 0-1 days for most of the bottles

